Weighted Model Counting Without Parameter Variables

### 2021-02-16 doc/paper3/paper.tex
Writing and editing

@@

\begin{lemma} \label{lemma:add_size}
  An ADD with $n$ variables has $\mathcal{O}(2^n)$ nodes.
\end{lemma}
\begin{proof}
  An ADD with $n$ variables can be, at most, a complete binary tree of height
  $n+1$ (as measured by the number of vertices). It would then have $2^0 + 2^1 +
  \dots + 2^n = 2^{n=1} = \mathcal{O}(2^n)$ nodes.
\end{proof}

\begin{lemma}
  Let $\phi$ be a conjunction of $n$ literals. Then the ADD representation of
  $[\phi]^p_q$ can be constructed in $\mathcal{O}(2^n)$ time for any $p, q \in
  \mathbb{R}$ such that $p \ne q$.
\end{lemma}
\begin{proof}
  The ADD representation of $\phi$ itself can be constructed with a sequence of
  $n-1$ calls to \texttt{apply} with one of the two operands in each call always
  a literal. The number of variables in the other operand then follows the
  sequence $1, 2, 3, \dots, n-1$. By \cref{lemma:add_size}, the numbers of nodes
  in the ADD representations of these operands is then $\mathcal{O}(2^1),
  \mathcal{O}(2^2), \dots, \mathcal{O}(2^{n-1})$. Since one of the operands is
  of constant size, the overall time complexity of all calls to \texttt{apply}
  is then
  \[
    \mathcal{O}(2^1) + \mathcal{O}(2^2) + \dots + \mathcal{O}(2^{n-1}) =
    \mathcal{O}(2^n).
  \]

  Let $\alpha$ be the ADD representation of $\phi$. Then the ADD representation
  of $[\phi]_q^p$ is $(p-q)\alpha + q$. As scalar operations can obviously be
  implemented in linear time, the overall complexity remains $\mathcal{O}(2^n)$.
\end{proof}
% TODO: will need to define scalar operations on ADDs

@@

\paragraph{Notes.}
\begin{itemize}
\item Apparently, the DPMC paper already shows that taking the first offered
  decomposition tree is best.
\item It is already well-known that WMC is FPT.
\end{itemize}

@@

\begin{algorithm}
  \caption{WMC instance transformation}
  \SetKwFunction{rename}{rename}
  \SetKwProg{Fn}{Function}{:}{}
  \KwData{an (old-format) WMC instance $(\phi, X_I, X_P, W)$}
  \KwResult{a (new-format) WMC instance $(\phi', \omega)$}
  $\phi' \gets \emptyset$\;
  $\omega \gets 1$\;
  let $d\colon X_P \to \mathbb{N}$ be defined as $v \mapsto |\{ u \in X_P \mid u
  \le v \}|$\;
  \ForEach{clause $c \in \phi$}{
    \uIf{$c \cap X_P = \{ v \}$ for some $v$ \textnormal{\textbf{and}} $W(v) \ne
      1$}{
      \uIf{$|c| = 1$}{$\omega \gets \omega \times W(v)$\;}
      \Else{
        $\phi' \gets \phi' \cup \left\{ \left[ \bigwedge_{l \in c \setminus
              \{v\}} \neg l \right]^{W(v)}_1 \right\}$\;
      }
    }
    \ElseIf{$\{v \mid \neg v \in c\} \cap X_P  = \emptyset$}{
      $\phi' \gets \phi' \cup \{ [c]^1_0 \}$\;
    }
  }
  \ForEach{indicator variable $v \in X_I$}{
    \If{$\{[v]_1^p, [\neg v]_1^q\} \subseteq \phi'$ for some $p$ and $q$}{
      $\phi' \gets \phi' \setminus \{ [v]_1^p, [\neg v]_1^q \} \cup \{ [v]_q^p
      \}$\;
    }
  }
  replace every variable $v$ in $\phi'$ with $\rename{$v$}$\;
  \Return{$(\phi', \omega)$}\;
  \Fn{\rename{$v$}}{
    $S \gets \{u \in X_P \mid u \le v\}$\;
    \lIf{$S = \emptyset$}{\Return{$v$}}
    \Return{$v - d(\max S)$}\;
  }
\end{algorithm}

\section{Parameterised Complexity of DPMC}

\paragraph{Notes.}
\begin{itemize}
\item Summary of results
  \begin{itemize}
  \item We establish DPMC inference as fixed-parameter tractable.
  \item We experimentally show that cw+DPMC is best on low-to-moderate treewidth
    instances, and cd06+c2d overtakes cw+DPMC on higher treewidth instances.
  \end{itemize}
\item By DPMC, we always mean DMC+lg.
\end{itemize}

\paragraph{TODO}
\begin{itemize}
\item do a literature search focused around these papers
\item use mathcal on V, L, E (and introduce them)
\item maybe use texttt for DPMC?
\item do I need to formally consider extending a pseudo-Boolean function to a
  bigger domain?
\item Define:
  \begin{itemize}
  \item an ADD as a DAG.
  \item What does it mean for an ADD to `have' variables? Maybe refer to
    pseudo-Boolean function sensitivity.
  \item Formal definition of a previous WMC instance (CNF, literal weight
    function) and the new definition (set of $2^{X} \to \mathbb{R}_{\ge 0}$
    pseudo-Boolean functions and a constant). Note that the constant idea is
    borrowed from \texttt{bklm16}.
  \item Boolean formula in CNF (perhaps this is too trivial to define),
  \item primal graph of a CNF formula (a.k.a.
    Gaifman/(variable) interaction/connectivity/clique/representing graph),
  \item Bayesian network (I already have a definition of these last two),
  \item moralisation of a Bayesian network (or of any DAG),
  \end{itemize}
\end{itemize}

@@

\begin{definition}[\cite{DBLP:conf/cp/DudekPV20}, verbatim but with one change]
  Let $X$ be a set of Boolean variables and $\phi$ be a CNF formula over $X$. A
  \emph{project-join tree} (PJT) of $\phi$ is a tuple $(T, r, \gamma, \pi)$
  where:
  \begin{itemize}
  \item $T$ is a tree with root $r \in \mathcal{V}(T)$,
  \item $\gamma\colon \mathcal{L}(T) \to \mathbb{R}^{2^X}$ maps leaves of $T$ to
    pseudo-Boolean functions, and
  \item $\pi\colon \mathcal{V}(T) \setminus \mathcal{L}(T) \to 2^X$ is a
    labelling function on internal nodes.
  \end{itemize}
  Moreover, $(T, r, \gamma, \pi)$ must satisfy the following two properties:
  \begin{enumerate}
  \item $\{\pi(n) : n \in \mathcal{V}(T) \setminus \mathcal{L}(T)\}$ is a
    partition of $X$, and
  \item for each internal node $n \in \mathcal{V}(T) \setminus \mathcal{L}(T)$,
    variable $x \in \pi(n)$, and clause $c \in \phi$ such that $x$ appears in
    $c$, the leaf node $\gamma^{-1}(c)$ must be a descendant of $n$ in $T$.
  \end{enumerate}
\end{definition}

Note that while Dudek et al. \cite{DBLP:conf/cp/DudekPV20} define $\gamma$ to be
a bijection between leaves of a PJT and clauses, our definition is more
general.

\begin{definition}[\cite{DBLP:conf/cp/DudekPV20}]
  \[
    \mathtt{Vars}(n) \coloneqq
    \begin{cases}
      \mathtt{Vars}(\gamma(n)) & \text{if } n \in \mathcal{L}(T) \\
      \left( \bigcup_{o \in \mathcal{C}(n)} \mathtt{Vars}(o) \right) \setminus
      \pi(n) & \text{if } n \not\in \mathcal{L}(T).
    \end{cases}
  \]
  \[
    \mathtt{size}(n) \coloneqq
    \begin{cases}
      |\mathtt{Vars}(n)| & \text{if } n \in \mathcal{L}(T) \\
      |\mathtt{Vars}(n) \cup \pi(n)| & \text{if } n \not\in \mathcal{L}(T).
    \end{cases}
  \]
  (Note that $\mathtt{size}(n)$ is the number of variables that can appear
  during the computation of $\delta(n)$ (excluding recursive calls).)
  The \emph{width} of a PJT $(T, r, \gamma, \pi)$ is
  $\mathtt{width}(T) \coloneqq \max_{n \in \mathcal{V}(T)} \mathtt{size}(n)$.
\end{definition}

@@

\begin{definition}[\cite{DBLP:journals/jct/RobertsonS84}, rephrased]
  A \emph{tree decomposition} of a graph $G$ is a pair $(T, \chi)$, where $T$ is
  a tree and $\chi\colon V(T) \to 2^{V(G)}$ is a labelling function, with the
  following properties:
  \begin{itemize}
  \item $\bigcup_{t \in V(T)} \chi(t) = V(G)$;
  \item for every edge $e \in E(G)$, there exists $t \in V(T)$ such that $e$ has
    both endpoints in $\chi(t)$;
  \item for all $t, t', t'' \in V(T)$, if $t'$ is on the path between $t$ and
    $t''$, then $\chi(t) \cap \chi(t'') \subseteq \chi(t')$.
  \end{itemize}
  The \emph{width} of tree decomposition $(T, \chi)$ is $\max_{t \in V(T)}
  |\chi(t)| - 1)$. The \emph{treewidth} of graph $G$ is the smallest $w$ such
  that $G$ has a tree decomposition of width $w$.
\end{definition}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node[circle,draw] {$\exists$}
    child {node (a) {$\vdots$}
      child {node[circle,draw] {$c_1$}}
    }
    child {node (b) {$\vdots$}
      child {node[circle,draw] {$\ast$}
        child {node {$\vdots$}
          child {node[circle,draw] {$c_i$}}
        }
      }
    }
    child {node (c) {$\vdots$}
      child {node[circle,draw] {$c_l$}}
    }
    ;
    \path (a) -- (b) node[midway] {$\cdots$};
    \path (b) -- (c) node[midway] {$\cdots$};
  \end{tikzpicture}
  \caption{An example PJT with clauses $c_1, \dots, c_i, \dots, c_l$ that all
    contain the same variable, its projection node $\exists$, and a node under
    consideration $\ast$.}
  \label{fig:example_pjt}
\end{figure}

@@

% NOTE: this is a rehash of equation (2) from the DPMC paper
\begin{definition} \label{def:abstraction}
  Let $(T, r, \gamma, \pi)$ be a PJT, and $X$ be the set of variables. The
  functionality of DPMC execution can be represented by $\delta(r)$, where
  $\delta\colon \mathcal{V}(T) \to \mathbb{R}^{2^X}$ is a recursive function
  defined as
  \begin{equation} \label{eq:execution}
    \delta(t) =
    \begin{cases}
      \gamma(t) & \text{if } t \in \mathcal{L}(T) \\
      \exists_{\pi(t)} \prod_{u \in \mathcal{C}(t)} \delta(u) & \text{otherwise.}
    \end{cases}
  \end{equation}
  The range of $\delta(r)$ then contains a single real number, i.e., the answer.
\end{definition}

@@

% \begin{lemma} % NOTE: this lemma is already proven above
%   Let $(S, \chi)$ be a tree decomposition, and $(T, r, \gamma, \pi)$ be its PJT
%   generated by Algorithm~2 in \cite{DBLP:conf/cp/DudekPV20}. For any $t \in T$,
%   the number of variables that appear in the ADDs within a PJT node is at
%   most $k+1$.
% \end{lemma}
% \begin{proof}
%   All such variables are introduced in a `clause' below and not projected in
%   descendants. Each such variable would be in the bag of each clause that
%   introduced it. By the tree decomposition properties, it will also be in the
%   bag of both the projection node and the current node because they are on the
%   path between the clause nodes (see \cref{fig:example_pjt}).
% \end{proof}

\begin{theorem}
  Let $(T, r, \gamma, \pi)$ be a PJT of width $k$. Then \textsc{DPMC} execution
  is fixed-parameter tractable with respect to the $k$. Specifically,
  \textsc{DPMC} execution time complexity is $\mathcal{O}(4^km(n+k))$, where $n
  = |\mathcal{L}(T)|$ is the number of clauses/leaves, and $m$ is the number of
  variables.
\end{theorem}
% TODO: start here. Use theorem 4 to show that the number of variables is
% bounded by primal treewidth
\begin{proof}
  \begin{itemize}
  \item Let us consider the overall complexity of all multiplications and
    projections throughout the recursive calls of $\delta$ in
    \cref{def:abstraction}. Clearly, the overall complexity is the sum of the
    complexity of operations performed within each call to $\delta$.
  \item Let $t \in \mathcal{V}(T)$ be an arbitrary vertex of $T$. If $t$ is a
    leaf,
  \item The total time taken by the DPMC execution stage is the sum of the time
    taken `inside' each PJT node. The number of such nodes is upper bounded by
    $m$, so the total time complexity of DPMC is $\mathcal{O}(4^km(n+k))$.
  \item Multiplying $m$ ADDs can then take up to $(m-1)(2^{k+1}-1)^2 =
    \mathcal{O}(m4^k)$ (since one multiplication takes up to $(2^{k+1}-1)^2$
    time and the and the result will have up to $2^{k+1} - 1$ nodes because the
    domain stays the same).
  \item Then, by \cref{lemma:basic_ops}, projecting $m$ variables will take
    $\mathcal{O}(4^km)$ time.
  \item In total, operations on a PJT node with $m$ children, $k$ variables, $n$ of
    which are projected takes $\mathcal{O}(4^k(m+n))$ time.
  \end{itemize}
\end{proof}

>>>>

\DeclareMathOperator{\im}{im}

@@

\begin{lemma} \label{lemma:add_size}
  An ADD with $n$ variables has $\mathcal{O}(2^n)$ nodes.
\end{lemma}
\begin{proof}
  An ADD with $n$ variables can be, at most, a complete binary tree of height
  $n+1$ (as measured by the number of vertices). It would then have $2^0 + 2^1 +
  \dots + 2^n = 2^{n+1} = \mathcal{O}(2^n)$ nodes.
\end{proof}

\begin{lemma} \label{lemma:clause_time}
  Let $\phi$ be a conjunction of $n$ literals. Then the ADD representation of
  $[\phi]^p_q$ can be constructed in $\mathcal{O}(2^n)$ time for any $p, q \in
  \mathbb{R}$ such that $p \ne q$.
\end{lemma}
\begin{proof}
  The ADD representation of $\phi$ itself can be constructed with a sequence of
  $n-1$ calls to \texttt{apply} with one of the two operands in each call always
  a literal. The number of variables in the other operand then follows the
  sequence $1, 2, 3, \dots, n-1$. By \cref{lemma:add_size}, the numbers of nodes
  in the ADD representations of these operands is then $\mathcal{O}(2^1),
  \mathcal{O}(2^2), \dots, \mathcal{O}(2^{n-1})$. Since one of the operands is
  of constant size, the overall time complexity of all calls to \texttt{apply}
  is then
  \[
    \mathcal{O}(2^1) + \mathcal{O}(2^2) + \dots + \mathcal{O}(2^{n-1}) =
    \mathcal{O}(2^n).
  \]

  Let $\alpha$ be the ADD representation of $\phi$. Then the ADD representation
  of $[\phi]_q^p$ is $(p-q)\alpha + q$. As scalar operations can obviously be
  implemented in linear time, the overall complexity remains $\mathcal{O}(2^n)$.
\end{proof}

% NOTE: weights on literals other than the positive literals that correspond to
% variables in $X_P$ are redundant as they either are equal to one or duplicate
% an already-defined weight.
\begin{definition}[old WMC instance]
  An old WMC instance is a tuple $(\phi, X_I, X_P, w)$, where $X_I$ is the set
  of indicator variables, $X_P$ is the set of parameter variables, $\phi$ is a
  propositional formula over $X_I \cup X_P$, and $w\colon X_P \to \mathbb{R}$ is
  the weight function.
\end{definition}

\begin{definition}[new WMC instance]
  A new WMC instance is a tuple $(F, X, \omega)$, where $X$ is the set of
  variables, $F$ is a set of pseudo-Boolean functions $2^X \to \mathbb{R}$, and
  $\omega \in \mathbb{R}$ is the scaling factor.
\end{definition}

@@

\paragraph{Notes.}
\begin{itemize}
\item Apparently, the \textsc{DPMC} paper already shows that taking the first
  offered decomposition tree is best.
\item It is already well-known that WMC is FPT.
\end{itemize}

@@

\begin{algorithm}
  \caption{WMC instance transformation}
  \SetKwFunction{rename}{rename}
  \SetKwProg{Fn}{Function}{:}{}
  \KwData{an (old-format) WMC instance $(\phi, X_I, X_P, w)$}
  \KwResult{a (new-format) WMC instance $(F, X, \omega)$}
  $F \gets \emptyset$\;
  $\omega \gets 1$\;
  let $d\colon X_P \to \mathbb{N}$ be defined as $v \mapsto |\{ u \in X_P \mid u
  \le v \}|$\;
  \ForEach{clause $c \in \phi$}{
    \uIf{$c \cap X_P = \{ v \}$ for some $v$ \textnormal{\textbf{and}} $w(v) \ne
      1$}{
      \uIf{$|c| = 1$}{$\omega \gets \omega \times w(v)$\;}
      \Else{
        $F \gets F \cup \left\{ \left[ \bigwedge_{l \in c \setminus
              \{v\}} \neg l \right]^{w(v)}_1 \right\}$\;
      }
    }
    \ElseIf{$\{v \mid \neg v \in c\} \cap X_P  = \emptyset$}{
      $F \gets F \cup \{ [c]^1_0 \}$\;
    }
  }
  \ForEach{indicator variable $v \in X_I$}{
    \If{$\{[v]_1^p, [\neg v]_1^q\} \subseteq F$ for some $p$ and $q$}{
      $F \gets F \setminus \{ [v]_1^p, [\neg v]_1^q \} \cup \{ [v]_q^p
      \}$\;
    }
  }
  replace every variable $v$ in $F$ with $\rename{$v$}$\;
  \Return{$(F, X_I, \omega)$}\;
  \Fn{\rename{$v$}}{
    $S \gets \{u \in X_P \mid u \le v\}$\;
    \lIf{$S = \emptyset$}{\Return{$v$}}
    \Return{$v - d(\max S)$}\;
  }
\end{algorithm}

\section{Parameterised Complexity of \textsc{DPMC}}

\paragraph{Notes.}
\begin{itemize}
\item Summary of results
  \begin{itemize}
  \item We establish \textsc{DPMC} inference as fixed-parameter tractable.
  \item We experimentally show that \textsc{DPMC} is best on low-to-moderate
    treewidth instances, and cd06+c2d overtakes on higher treewidth instances.
  \end{itemize}
\item By \textsc{DPMC}, we always mean DMC+lg.
\end{itemize}

\paragraph{TODO}
\begin{itemize}
\item do a literature search focused around these papers
\item do I need to formally consider extending a pseudo-Boolean function to a
  bigger domain?
\item Introduce all the notation surrounding graphs ($\mathcal{V}$,
  $\mathcal{L}$, $\mathcal{C}$, etc.) (and be consistent with its usage)
\item Define:
  \begin{itemize}
  \item scalar operations on ADDs
  \item an ADD as a DAG.
  \item What does it mean for an ADD to `have' variables? Maybe refer to
    pseudo-Boolean function sensitivity.
  \item Formal definition of a previous WMC instance (CNF, literal weight
    function) and the new definition (set of $2^{X} \to \mathbb{R}_{\ge 0}$
    pseudo-Boolean functions and a constant). Note that the constant idea is
    borrowed from \texttt{bklm16}.
  \item Boolean formula in CNF (perhaps this is too trivial to define),
  \item primal graph of a CNF formula (a.k.a.
    Gaifman/(variable) interaction/connectivity/clique/representing graph),
  \item Bayesian network (I already have a definition of these last two),
  \item moralisation of a Bayesian network (or of any DAG),
  \end{itemize}
\end{itemize}

@@

\begin{definition}[\cite{DBLP:conf/cp/DudekPV20}, with some changes]
  Let $X$ be a set of Boolean variables, and $F$ be a set of pseudo-Boolean
  functions $2^X \to \mathbb{R}$. A \emph{project-join tree} (PJT) of $F$ is a
  tuple $(T, r, \gamma, \pi)$ where:
  \begin{itemize}
  \item $T$ is a tree with root $r \in \mathcal{V}(T)$,
  \item $\gamma\colon \mathcal{L}(T) \to F$ is a bijection between the leaves of
    $T$ and the pseudo-Boolean functions in $F$, and
  \item $\pi\colon \mathcal{V}(T) \setminus \mathcal{L}(T) \to 2^X$ is a
    labelling function on internal nodes.
  \end{itemize}
  Moreover, $(T, r, \gamma, \pi)$ must satisfy the following two properties:
  \begin{enumerate}
  \item $\{\pi(n) : n \in \mathcal{V}(T) \setminus \mathcal{L}(T)\}$ is a
    partition of $X$, and
  \item for each internal node $n \in \mathcal{V}(T) \setminus \mathcal{L}(T)$,
    variable $x \in \pi(n)$, and `clause' $c \in \im \gamma$ such that $x$
    appears in $c$, the leaf node $\gamma^{-1}(c)$ must be a descendant of $n$
    in $T$.
  \end{enumerate}
\end{definition}

\begin{definition}[\cite{DBLP:conf/cp/DudekPV20}] \label{def:size}
  \[
    \mathtt{Vars}(n) \coloneqq
    \begin{cases}
      \mathtt{Vars}(\gamma(n)) & \text{if } n \in \mathcal{L}(T) \\
      \left( \bigcup_{o \in \mathcal{C}(n)} \mathtt{Vars}(o) \right) \setminus
      \pi(n) & \text{if } n \not\in \mathcal{L}(T).
    \end{cases}
  \]
  \[
    \mathtt{size}(n) \coloneqq
    \begin{cases}
      |\mathtt{Vars}(n)| & \text{if } n \in \mathcal{L}(T) \\
      |\mathtt{Vars}(n) \cup \pi(n)| & \text{if } n \not\in \mathcal{L}(T).
    \end{cases}
  \]
  (Note that $\mathtt{size}(n)$ is the number of variables that can appear
  during the computation of $\delta(n)$ (excluding recursive calls).)
  The \emph{width} of a PJT $(T, r, \gamma, \pi)$ is
  $\mathtt{width}(T) \coloneqq \max_{n \in \mathcal{V}(T)} \mathtt{size}(n)$.
\end{definition}

@@

\begin{definition}[\cite{DBLP:journals/jct/RobertsonS84}, rephrased]
  A \emph{tree decomposition} of a graph $G$ is a pair $(T, \chi)$, where $T$ is
  a tree and $\chi\colon \mathcal{V}(T) \to 2^{\mathcal{V}(G)}$ is a labelling
  function, with the following properties:
  \begin{itemize}
  \item $\bigcup_{t \in \mathcal{V}(T)} \chi(t) = \mathcal{V}(G)$;
  \item for every edge $e \in \mathcal{E}(G)$, there exists $t \in
    \mathcal{V}(T)$ such that $e$ has both endpoints in $\chi(t)$;
  \item for all $t, t', t'' \in \mathcal{V}(T)$, if $t'$ is on the path between
    $t$ and $t''$, then $\chi(t) \cap \chi(t'') \subseteq \chi(t')$.
  \end{itemize}
  The \emph{width} of tree decomposition $(T, \chi)$ is $\max_{t \in
    \mathcal{V}(T)} |\chi(t)| - 1)$. The \emph{treewidth} of graph $G$ is the
  smallest $w$ such that $G$ has a tree decomposition of width $w$.
\end{definition}

@@

% NOTE: this is a rehash of equation (2) from the DPMC paper
\begin{definition} \label{def:abstraction}
  Let $(T, r, \gamma, \pi)$ be a PJT, and $X$ be the set of variables. The
  functionality of \textsc{DPMC} execution can be represented by $\delta(r)$,
  where $\delta\colon \mathcal{V}(T) \to \mathbb{R}^{2^X}$ is a recursive
  function defined as
  \begin{equation} \label{eq:execution}
    \delta(t) =
    \begin{cases}
      \gamma(t) & \text{if } t \in \mathcal{L}(T) \\
      \exists_{\pi(t)} \prod_{u \in \mathcal{C}(t)} \delta(u) & \text{otherwise.}
    \end{cases}
  \end{equation}
  The range of $\delta(r)$ then contains a single real number, i.e., the answer.
\end{definition}

@@

\begin{theorem}
  Let $(T, r, \gamma, \pi)$ be a PJT of width $k$. Then \textsc{DPMC} execution
  is fixed-parameter tractable with respect to the $k$. Specifically,
  \textsc{DPMC} execution time complexity is $\mathcal{O}(4^knm)$, where $n
  = |\mathcal{L}(T)|$ is the number of clauses/leaves, and $m$ is the number of
  variables.
\end{theorem}
% TODO: Use theorem 4 to show that the number of variables is
% bounded by primal treewidth
\begin{proof}
  \begin{itemize}
  \item Let us consider the overall complexity of all multiplications and
    projections throughout the recursive calls of $\delta$ in
    \cref{def:abstraction}. Clearly, the overall complexity is the sum of the
    complexity of operations performed within each call to $\delta$.
  \item Let $t \in \mathcal{L}(T)$ be a leaf. Note that
    $|\mathtt{Vars}(\gamma(t))| \le k$ by \cref{def:size}. Assuming that clauses
    have no repeating variables, it takes $\mathcal{O}(2^k)$ time to construct
    $\delta(t)$ by \cref{lemma:clause_time} and $\mathcal{O}(2^kn)$ time to
    construct all leaves of the PJT.
  \item Now let $t \in \mathcal{V}(T) \setminus \mathcal{L}(T)$ be an internal
    vertex, in which case we perform $|\mathcal{C}(t)|-1$ multiplications and
    $|\pi(t)|$ projections. Note that $|\mathcal{C}(t)| \le |\mathcal{L}(T)| =
    n$, $|\pi(t)| \le k$, and the ADDs involved can have up to $k$ variables.
    This means that each ADD has $\mathcal{O}(2^k)$ nodes (regardless of whether
    it comes from $\mathcal{C}(t)$ or from multiplications). Each such
    multiplication then takes $\mathcal{O}(4^k)$ time by \cref{lemma:basic_ops},
    and so all multiplications will take $\mathcal{O}(4^kn)$ time for $t$ and
    $\mathcal{O}(4^knm)$ time across all $t \in \mathcal{V}(T) \setminus
    \mathcal{L}(T)$ since the number of vertices in the PJT is bounded by the
    number of variables.
  \item Each variable is projected exactly once and is always projected from
    an ADD with at most $\mathcal{O}(2^k)$ nodes. Thus, the time complexity of
    projecting all variables is $\mathcal{O}(4^km)$.
  \item In total, we get $\mathcal{O}(2^kn)$ time for the leaves,
    $\mathcal{O}(4^knm)$ for multiplications, and $\mathcal{O}(4^km)$ for
    projections, resulting in $\mathcal{O}(4^knm)$ time in total.
  \end{itemize}
\end{proof}
